---
title: "Data Cleaning - Engagement, Espalier & Surveys"
author: "Xufei Lang"
output: html_document
date: "2026-01-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Download and install ODBC driver through this link: 
https://sfc-repo.snowflakecomputing.com/odbc/win64/3.14.0/snowflake64_odbc-3.14.0.msi


## Install Packages.
```{r warning=FALSE message=FALSE}
# install.packages(c("DBI", "dplyr","dbplyr","odbc"))  # remove the # sign to install packages.
# install.packages("lubridate")
library(DBI)
library(dplyr)
library(dbplyr)
library(odbc)
library(lubridate)
library(stringr)
library(tidyr)
```


## Connect to Snowflake.
```{r}
con <- dbConnect(
  odbc::odbc(),
  Driver     = "SnowflakeDSIIDriver",
  Server     = "CALIX.snowflakecomputing.com",
  Database   = "DOWNSTREAM_DATA_HUB",
  Schema     = "GA_TECH_ML",
  Warehouse  = "GA_TECH_ML_WH",
  Role       = "FUNC_GA_TECH_ML",
  Authenticator = "externalbrowser",
  UID      = "",    # replace with your Firstname.Lastname@CALIX.COM
  PWD      = ""    # replace with your password
)

# Test
dbGetQuery(con, "select current_role(), current_warehouse(), current_database(), current_schema()")
```


#################### Engagement(comm,conn,virt) ####################

## Pull tables from Snowflake.
```{r}
accounts <- dbGetQuery(con,"select * from accounts")   # change the table names as needed.
community_activity <- dbGetQuery(con,"select * from community_activity")
connexions_attendance <- dbGetQuery(con,"select * from connexions_attendance")
virtual_engagement_attendance <- dbGetQuery(con,"select * from virtual_engagement_attendance")
```


## Prepare tables.
```{r}
# Accounts table
account_id <- accounts %>% select(ACCOUNT_ID)
account_id_list <- accounts$ACCOUNT_ID
head(account_id)
dim(account_id)
```


```{r}
# Generate Normalized percentage weights table. 
weights_table <- function(start_year = 2020, end_year = 2025, r = 0.5) {
  years <- start_year:end_year
  cols <- lapply(start_year:end_year, function(sy) {
    yrs_win <- sy:end_year
    age <- end_year - yrs_win
    w_raw <- r ^ age
    w <- w_raw / sum(w_raw)
    col <- setNames(rep(0, length(years)), years)
    col[as.character(yrs_win)] <- round(100 * w, 1)
    col
  })
  df <- data.frame(
    YEAR = years,
    do.call(cbind, cols),
    check.names = FALSE
  )
  names(df)[-1] <- sprintf("weight_%02d_%02d", start_year:end_year, end_year)
  rownames(df) <- NULL
  df
}

# Example:
weights <- weights_table(2020, 2025, r=0.5)
weights
```


```{r}
# Community_activity table
head(community_activity)
dim(community_activity)
comm <- community_activity %>%
  filter(ACCOUNT_ID %in% account_id_list) %>%
  select(ACCOUNT_ID, ID, CREATED_DATE) %>%
  distinct(ID, .keep_all=TRUE) %>%
  mutate(YEAR = year(CREATED_DATE)) %>%
  filter(YEAR >= 2020) %>% # With geometric decay r=0.5, data older than ~5 years contributes almost nothing.
  group_by(ACCOUNT_ID,YEAR) %>%
  summarize(Comm_Cnt = n_distinct(ID), .groups='drop')
head(comm)
dim(comm)
sort(unique(comm$YEAR))
```

```{r}
# Apply normalized percentage weights to Community_activity table 
comm_cnt <- comm %>%
  left_join(weights[,c("YEAR","weight_2020_2025")], by="YEAR") %>%
  mutate(weighted_value = Comm_Cnt*weight_2020_2025/100) %>%
  group_by(ACCOUNT_ID) %>%
  summarize(Community_Cnt = sum(weighted_value, na.rm=TRUE), .groups="drop")
comm_cnt
```


```{r}
# Connexions_attendance table
head(connexions_attendance)
dim(connexions_attendance)
conn <- connexions_attendance %>%
  filter(ACCOUNT_ID %in% account_id_list) %>%
  filter(str_detect(CAMPAIGN_MEMBER_STATUS, regex("Attended", ignore_case=TRUE))) %>%
  select(ACCOUNT_ID, CAMPAIGN_ID, FIRST_RESPONDED_AT_DATE) %>%
  mutate(YEAR = year(FIRST_RESPONDED_AT_DATE)) %>%
  filter(YEAR >= 2020) %>% # With geometric decay r=0.5, data older than ~5 years contributes almost nothing.
  group_by(ACCOUNT_ID,YEAR) %>%
  summarize(Conn_Cnt = n_distinct(CAMPAIGN_ID), .groups='drop')
head(conn)
dim(conn)
sort(unique(conn$YEAR))
```

```{r}
# Apply normalized percentage weights to Community_activity table 
conn_cnt <- conn %>%
  left_join(weights[,c("YEAR","weight_2020_2025")], by="YEAR") %>%
  mutate(weighted_value = Conn_Cnt*weight_2020_2025/100) %>%
  group_by(ACCOUNT_ID) %>%
  summarize(Connexion_Cnt = sum(weighted_value, na.rm=TRUE), .groups="drop")
conn_cnt
```


```{r}
# Virtual_engagement_attendance table
head(virtual_engagement_attendance)
dim(virtual_engagement_attendance)
virt <- virtual_engagement_attendance %>%
  filter(ACCOUNT_ID %in% account_id_list) %>%
  filter(str_detect(CAMPAIGN_MEMBER_STATUS, regex("Attended", ignore_case=TRUE))) %>%
  select(ACCOUNT_ID, CAMPAIGN_ID, FIRST_RESPONDED_AT_DATE) %>%
  mutate(YEAR = year(FIRST_RESPONDED_AT_DATE)) %>%
  filter(YEAR >= 2020) %>% # With geometric decay r=0.5, data older than ~5 years contributes almost nothing.
  group_by(ACCOUNT_ID,YEAR) %>%
  summarize(Virt_Cnt = n_distinct(CAMPAIGN_ID), .groups='drop')
head(virt)
dim(virt)
sort(unique(virt$YEAR))
```

```{r}
# Apply normalized percentage weights to Community_activity table 
virt_cnt <- virt %>%
  left_join(weights[,c("YEAR","weight_2020_2025")], by="YEAR") %>%
  mutate(weighted_value = Virt_Cnt*weight_2020_2025/100) %>%
  group_by(ACCOUNT_ID) %>%
  summarize(Virtual_Cnt = sum(weighted_value, na.rm=TRUE), .groups="drop")
virt_cnt
```


## Join all three tables together to create a new dataframe, name it as Engagement
```{r}
Engagement <- account_id %>%
  left_join(comm_cnt, by="ACCOUNT_ID") %>%
  left_join(conn_cnt, by="ACCOUNT_ID") %>%
  left_join(virt_cnt, by="ACCOUNT_ID") %>%
  mutate(across(everything(), ~ replace_na(.x, 0))) %>%
  mutate(Log_Community_Cnt = round(log10(Community_Cnt+1),3)) %>%
  mutate(Log_Virtual_Cnt = round(log10(Virtual_Cnt+1),3)) %>%
  select(ACCOUNT_ID, Community_Cnt, Log_Community_Cnt, Connexion_Cnt, Virtual_Cnt, Log_Virtual_Cnt) %>%
  arrange(ACCOUNT_ID)
head(Engagement)
dim(Engagement)
```


## Visualize the final table. 
```{r}
hist(Engagement$Log_Community_Cnt)
hist(Engagement$Log_Virtual_Cnt)
hist(Engagement$Connexion_Cnt)
```


## Save table as csv file.
```{r}
write.csv(Engagement, "C:/Users/xlang/Downloads/Data/From RStudio/Engagement(comm,conn,virt).csv", row.names=FALSE)
```



#################### Espalier ####################
Note: Espalier from Sophie's file was the final version.

## Pull Espalier_customer_journey table from Snowflake.
```{r}
espalier_customer_journey <- dbGetQuery(con,"select * from espalier_customer_journey")
```

## Prepare table.
```{r}
head(espalier_customer_journey)
dim(espalier_customer_journey)
Espalier <- espalier_customer_journey %>%
  filter(SCORES_IS_LATEST_FLAG == TRUE) %>%
  filter(ACCOUNT_ID %in% account_id_list) %>%
  select(ACCOUNT_ID, CUSTOMER_JOURNEY_WEIGHTED_SCORE, SERVES_SMALL_BUSINESS, SERVES_MIDSIZE_BUSINESS, SERVES_LARGE_BUSINESS) %>%
  mutate(across(everything(), ~ replace_na(.x, FALSE))) %>%
  arrange(ACCOUNT_ID)
head(Espalier)  
dim(Espalier)
```


## Visualize table.
```{r}
hist(Espalier$CUSTOMER_JOURNEY_WEIGHTED_SCORE)
barplot(table(Espalier$SERVES_SMALL_BUSINESS), main="SERVES_SMALL_BUSINESS")
barplot(table(Espalier$SERVES_MIDSIZE_BUSINESS), main="SERVES_MIDSIZE_BUSINESS")
barplot(table(Espalier$SERVES_LARGE_BUSINESS), main="SERVES_LARGE_BUSINESS")
```


## Save table as csv file.
```{r}
write.csv(Espalier, "C:/Users/xlang/Downloads/Data/From RStudio/Espalier.csv", row.names=FALSE)
```



#################### Surveys ####################

## Pull Surveys table from Snowflake.
```{r}
surveys <- dbGetQuery(con,"select * from surveys")
```


## Prepare table.
```{r}
head(surveys)
dim(surveys)
Surveys <- surveys %>%
  filter(ACCOUNT_ID %in% account_id_list) %>%
  group_by(ACCOUNT_ID) %>%
  filter(SURVEY_CREATED_AT_DATE == max(SURVEY_CREATED_AT_DATE)) %>%
  mutate(NPS_GROUP_SCORE = case_when(NPS_GROUP == "promoter" ~ 1,
                                     NPS_GROUP == "passive" ~ 0,
                                     NPS_GROUP == "detractor" ~ -1)) %>%
  arrange(ACCOUNT_ID)
head(Surveys)  
dim(Surveys)
```


## Save table as csv file.
```{r}
write.csv(Surveys, "C:/Users/xlang/Downloads/Data/From RStudio/Surveys.csv", row.names=FALSE)
```

