---
title: "EDA(Exploratory Data Analysis) Insights"
output:
  pdf_document: default
  html_document: default
---

# How AIRBNB price changes on monthly basis

The aim of this EDA is to give an overview how AIRBNB price rate changes on month to month basis and Boston city was taken for reference. 

```{r}
#installing packages
library("colorspace")
library("ggplot2")
library("tidyverse")
library("dplyr")
library("data.table")

#Setting the directort for the data set
setwd("/Users/pooja/Desktop/GT_College/08_MGT_6203/Group_Project/eda_work")

#Reading Boston Calendar data set from Sep-2022 to June-2022
boston_whole_dataset <- read.csv("boston_whole.csv")
head(boston_whole_dataset, 5)

#Summing of the price monthwise 
boston_whole_dataset1 = aggregate(list(price=boston_whole_dataset$price),by=list(month_details=cut(as.POSIXct(boston_whole_dataset$date),"month")),sum)
boston_whole_dataset1

#Renaming the row names to actual month names
boston_whole_dataset1$month_details <- c("Sep","Oct","Nov","Dec","Jan","Feb","Mar","Apr",
                                     "May","June")
boston_whole_dataset1

#plotting the income vs month
boston_whole_dataset1 %>% 
  group_by(price,month_details) %>%
  ggplot(aes(y=price, x=month_details)) + 
  geom_col(fill = "blue", alpha = 0.5)+
  theme_grey()+
  theme(panel.grid.minor = element_blank())+
  labs(title = "How AIRBNB price changes on monthly basis",
       x = "Month", y = "Price", fill = "Room Type")

```
### Observations:
- We can see that October month where the price was high. 
- We assume that during October month the COVID-19 risk was little low due to which there might be more demand in October month and price is high


# Which room type is more popular

The aim of this EDA is to give an overview which room type has better review rating as well as cost effective 

```{r}
#Reading the Dallas data set
dallas_oct_data <- read.csv("dallas_eda.csv")
head(dallas_oct_data,5)

#checking the column names
colnames(dallas_oct_data)

#Plotting the graph between price, average score rating and room type
dallas_oct_data %>%
  group_by(room_type) %>%
  summarise(avg_price = mean(price), avg_review= mean(review_scores_rating)) %>%
  ggplot(aes(x=room_type, fill=avg_price, y = avg_review))+
  geom_col(color = "White")+
  theme_grey()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())+
  labs(title = "Which room types are most popular?",
       x = "Room Type", y = "Average Reviews", fill = "Price")
```

### Observations:
- Entire home is most epxensive where shared room is cheaper which is making sense. 
- Both entire home as well as provate room has better review rating. 
